---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  dev = "png",
  dev.args = list(type = "cairo-png"),
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# fmesher: Triangle Meshes and Other Geometry Tools

<!-- badges: start -->
[![CRAN status](http://www.r-pkg.org/badges/version-last-release/fmesher)](https://cran.r-project.org/package=fmesher)
[![inlabru-org r-universe status](https://inlabru-org.r-universe.dev/badges/fmesher)](https://inlabru-org.r-universe.dev)
[![R build status](https://github.com/inlabru-org/fmesher/workflows/R-CMD-check/badge.svg)](https://github.com/inlabru-org/fmesher/actions)
[![test-coverage](https://github.com/inlabru-org/fmesher/workflows/test-coverage/badge.svg)](https://github.com/inlabru-org/fmesher/actions)
[![Codecov test coverage](https://codecov.io/gh/inlabru-org/fmesher/branch/devel/graph/badge.svg)](https://app.codecov.io/gh/inlabru-org/fmesher?branch=devel)
<!-- badges: end -->

Generate planar and spherical triangle meshes,
compute finite element calculations for 1- and 2-dimensional flat and curved
manifolds with associated basis function spaces, methods for lines and
polygons, and transparent handling of coordinate reference systems and
coordinate transformation, including 'sf' and 'sp' geometries.
The core 'fmesher' library code was originally part of the
[`INLA`](https://www.r-inla.org) package, and also distributed in the
[EUSTACE Horizon 2020 project](https://github.com/eustace-data/eustace-ambitious).
The expanded crs/CRS support started as an add-on feature of
[`inlabru`](https://inlabru-org.github.io/inlabru/).
    
## Installation

You can install the current [CRAN](https://CRAN.R-project.org) version of `fmesher`:
```{r cran-installation, eval = FALSE}
install.packages("fmesher")
```

You can install the latest stable release of `fmesher` from
[GitHub](https://github.com/):
``` r
# install.packages("remotes")
remotes::install_github("inlabru-org/fmesher", ref = "stable")
```

And the development version with:
``` r
# install.packages("remotes")
remotes::install_github("inlabru-org/fmesher")
```
or track the development version builds via [inlabru-org.r-universe.dev](https://inlabru-org.r-universe.dev/builds):
``` r
# Enable universe(s) by inlabru-org
options(repos = c(
  inlabruorg = "https://inlabru-org.r-universe.dev",
  INLA = "https://inla.r-inla-download.org/R/testing",
  CRAN = "https://cloud.r-project.org"
))

install.packages("fmesher")
```

To install and run `fmesher` in full debug mode (this is quite an experience!),
use
``` r
# install.packages("pkgbuild")
source("https://raw.githubusercontent.com/inlabru-org/fmesher/devel/misc/build.R")
fmesher_install(repo = "inlabru-org/fmesher", debug = TRUE)
```



## Online documentation

[https://inlabru-org.github.io/fmesher/](https://inlabru-org.github.io/fmesher/)

## Examples

### 2D triangular meshes

Includes a port of inla mesh `inla.mesh.create` (as `fm_rcdt_2d_inla()`) and
`inla.mesh.2d` interfaces.
```{r example2}
suppressPackageStartupMessages(library(fmesher))
suppressPackageStartupMessages(library(ggplot2))

(mesh <- fm_mesh_2d_inla(
  boundary = fm_extensions(cbind(0, 0), convex = c(1, 1.5)),
  max.edge = c(0.5, 1)
))
```
```{r example2-plot}
ggplot() +
  geom_fm(data = mesh) +
  theme_minimal()
```

### 1D B-spline function spaces

```{r example1}
(mesh <- fm_mesh_1d(c(1, 2, 3, 4, 6),
  boundary = c("neumann", "free"),
  degree = 2
))
```
```{r example1-plot}
ggplot() +
  geom_fm(data = mesh, xlim = c(0, 7))
```

### Extended helper methods for CRS handling

The package provides methods `fm_crs()` and `fm_CRS()` for extracting CRS
information from `sf` and `sp` objects and automatically converts to the
desired output format. The `fm_transform()` wrapper similarly handles a variety
of objects, as well as special handling for converting between spheres and globes
of different radii, e.g. used to map between the Earth and a unit radius sphere
uses as a model of the Earth.

```{r crs}
# longlat for a spherical version of the Earth
print(fm_crs("longlat_globe"))

# longlat for a sphere of radius 1m
print(fm_crs("longlat_norm"))

# A sphere of radius 1m
print(fm_crs("sphere"))
```
