---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  dev = "png",
  dev.args = list(type = "cairo-png"),
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# fmesher: Triangle Meshes and Other Geometry Tools

<!-- badges: start -->
[![inlabru status badge](https://inlabru-org.r-universe.dev/badges/fmesher)](https://inlabru-org.r-universe.dev)
[![R build status](https://github.com/inlabru-org/fmesher/workflows/R-CMD-check/badge.svg)](https://github.com/inlabru-org/fmesher/actions)
[![test-coverage](https://github.com/inlabru-org/fmesher/workflows/test-coverage/badge.svg)](https://github.com/inlabru-org/fmesher/actions)
[![Codecov test coverage](https://codecov.io/gh/inlabru-org/fmesher/branch/devel/graph/badge.svg)](https://app.codecov.io/gh/inlabru-org/fmesher?branch=devel)
<!-- badges: end -->

Functions that generate planar and spherical triangle meshes,
computes finite element calculations for 1- and 2-dimensional flat and curved
manifolds, functions for lines and polygons, and transparent handling of
crs information and coordinate transformation, including and interoperability
with `sf` and `sp` geometries. Experimental package that will take over the
relevant subset of features from the [INLA](https://www.r-inla.org) and
[inlabru](https://inlabru-org.github.io/inlabru/) packages, as well as extend
them.
    
## Installation

You can install the latest stable release of `fmesher` from
[GitHub](https://github.com/):
``` r
# install.packages("remotes")
remotes::install_github("inlabru-org/fmesher", ref = "stable")
```

And the development version with:
``` r
# install.packages("remotes")
remotes::install_github("inlabru-org/fmesher")
```
or track the development version builds via [inlabru-org.r-universe.dev](https://inlabru-org.r-universe.dev/builds):
``` r
# Enable universe(s) by inlabru-org
options(repos = c(
  inlabruorg = "https://inlabru-org.r-universe.dev",
  INLA = "https://inla.r-inla-download.org/R/testing",
  CRAN = "https://cloud.r-project.org"
))

install.packages("fmesher")
```
Note that you need the `feature/fmesher` branch of `inlabru` until that is merged
into the main development branch:
``` r
# install.packages("remotes")
remotes::install_github("inlabru-org/inlabru@feature/fmesher")
```

To install and run `fmesher` in full debug mode (this is quite an experience!),
use
``` r
# install.packages("pkgbuild")
source("https://raw.githubusercontent.com/inlabru-org/fmesher/devel/misc/build.R")
fmesher_install(repo = "inlabru-org/fmesher", debug = TRUE)
```



## Online documentation

[https://inlabru-org.github.io/fmesher/](https://inlabru-org.github.io/fmesher/)

## Examples

### 2D triangular meshes

Includes a port of inla mesh `inla.mesh.create` (as `fm_rcdt_2d_inla()`) and
`inla.mesh.2d` interfaces.
```{r example2}
suppressPackageStartupMessages(library(fmesher))
suppressPackageStartupMessages(library(ggplot2))

(mesh <- fm_mesh_2d_inla(
  boundary = fm_extensions(cbind(0, 0), convex = c(1, 1.5)),
  max.edge = c(0.5, 1)
))
```
```{r example2-plot}
ggplot() +
  geom_fm(data = mesh) +
  theme_minimal()
```

### 1D B-spline function spaces

```{r example1}
(mesh <- fm_mesh_1d(c(1, 2, 3, 4, 6),
  boundary = c("neumann", "free"),
  degree = 2
))
```
```{r example1-plot}
ggplot() +
  geom_fm(data = mesh, xlim = c(0, 7))
```

### Extended helper methods for CRS handling

The package provides methods `fm_crs()` and `fm_CRS()` for extracting CRS
information from `sf` and `sp` objects and automatically converts to the
desired output format. The `fm_transform()` wrapper similarly handles a variety
of objects, as well as special handling for converting between spheres and globes
of different radii, e.g. used to map between the Earth and a unit radius sphere
uses as a model of the Earth.

```{r crs}
# longlat for a spherical version of the Earth
print(fm_crs("longlat_globe"))

# longlat for a sphere of radius 1m
print(fm_crs("longlat_norm"))

# A sphere of radius 1m
print(fm_crs("sphere"))
```
