---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  dev = "png",
  dev.args = list(type = "cairo-png"),
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# fmesher: Triangle Meshes and Other Geometry Tools

<!-- badges: start -->
[![CRAN status](http://www.r-pkg.org/badges/version-last-release/fmesher)](https://cran.r-project.org/package=fmesher)
[![inlabru-org r-universe status](https://inlabru-org.r-universe.dev/badges/fmesher)](https://inlabru-org.r-universe.dev)
[![R build status](https://github.com/inlabru-org/fmesher/workflows/R-CMD-check/badge.svg)](https://github.com/inlabru-org/fmesher/actions)
[![test-coverage](https://github.com/inlabru-org/fmesher/workflows/test-coverage/badge.svg)](https://github.com/inlabru-org/fmesher/actions)
[![Codecov test coverage](https://codecov.io/gh/inlabru-org/fmesher/branch/devel/graph/badge.svg)](https://app.codecov.io/gh/inlabru-org/fmesher?branch=devel)
<!-- badges: end -->

Generate planar and spherical triangle meshes,
compute finite element calculations for 1- and 2-dimensional flat and curved
manifolds with associated basis function spaces, methods for lines and
polygons, and transparent handling of coordinate reference systems and
coordinate transformation, including 'sf' and 'sp' geometries.
The core 'fmesher' library code was originally part of the
[`INLA`](https://www.r-inla.org) package, and also distributed in the
[EUSTACE Horizon 2020 project](https://github.com/eustace-data/eustace-ambitious),
and implements parts of "Triangulations and Applications" by
[Hjelle and Dæhlen (2006)](https://doi.org/10.1007/3-540-33261-8).
The expanded crs/CRS support started as an add-on feature of
[`inlabru`](https://inlabru-org.github.io/inlabru/).
    
## Installation

You can install the current [CRAN version](https://cran.r-project.org/package=fmesher) version of fmesher:
```{r cran-installation, eval = FALSE}
install.packages("fmesher")
```

### Installation using [pak](https://pak.r-lib.org/)

You can install the latest bugfix release of fmesher from [GitHub](https://github.com/inlabru-org/fmesher) with:
```{r gh-bugfix-installation-pak, eval = FALSE}
# install.packages("pak")
pak::pkg_install("inlabru-org/fmesher@stable")
```

You can install the development version of inlabru from [GitHub](https://github.com/inlabru-org/fmesher) with
```{r gh-installation-pak, eval = FALSE}
pak::pkg_install("inlabru-org/fmesher@devel")
```
or track the development version builds via [inlabru-org.r-universe.dev](https://inlabru-org.r-universe.dev/builds):
```{r universe-installation-pak, eval = FALSE}
# Enable universe(s) by inlabru-org
pak::repo_add(inlabruorg = "https://inlabru-org.r-universe.dev")
pak::pkg_install("fmesher")
```
This will pick the r-universe version if it is more recent than the CRAN version.

To install and run `fmesher` in full debug mode (this is quite an experience!),
use
``` r
# install.packages("pkgbuild")
source("https://raw.githubusercontent.com/inlabru-org/fmesher/devel/misc/build.R")
fmesher_install(repo = "inlabru-org/fmesher", debug = TRUE)
```

### Installation using `remotes`

You can install the latest bugfix release of fmesher from [GitHub](https://github.com/inlabru-org/fmesher) with:
```{r gh-bugfix-installation-remotes, eval = FALSE}
# install.packages("remotes")
remotes::install_github("inlabru-org/fmesher", ref = "stable")
```

You can install the development version of fmesher from [GitHub](https://github.com/inlabru-org/fmesher) with
```{r gh-installation-remotes, eval = FALSE}
remotes::install_github("inlabru-org/fmesher")
```
or track the development version builds via [inlabru-org.r-universe.dev](https://inlabru-org.r-universe.dev/builds):
```{r universe-installation-remotes, eval = FALSE}
# Enable universe(s) by inlabru-org
options(repos = c(
  inlabruorg = "https://inlabru-org.r-universe.dev",
  getOption("repos")
))
install.packages("fmesher")
```

## Online documentation

[https://inlabru-org.github.io/fmesher/](https://inlabru-org.github.io/fmesher/)

## Examples

### 2D triangular meshes

Includes a port of inla mesh `inla.mesh.create` (as `fm_rcdt_2d_inla()`) and
`inla.mesh.2d` interfaces.
```{r example2}
suppressPackageStartupMessages(library(fmesher))
suppressPackageStartupMessages(library(ggplot2))

(mesh <- fm_mesh_2d_inla(
  boundary = fm_extensions(cbind(0, 0), convex = c(1, 1.5)),
  max.edge = c(0.5, 1)
))
```
```{r example2-plot,fig.alt="2D triangular mesh"}
ggplot() +
  geom_fm(data = mesh) +
  theme_minimal()
```

### 1D B-spline function spaces

```{r example1}
(mesh <- fm_mesh_1d(c(1, 2, 3, 4, 6),
  boundary = c("neumann", "free"),
  degree = 2
))
```
```{r example1-plot,fig.alt="1D B-spline function space"}
ggplot() +
  geom_fm(data = mesh, xlim = c(0, 7))
```

### Extended helper methods for CRS handling

The package provides methods `fm_crs()` and `fm_CRS()` for extracting CRS
information from `sf` and `sp` objects and automatically converts to the
desired output format. The `fm_transform()` wrapper similarly handles a variety
of objects, as well as special handling for converting between spheres and globes
of different radii, e.g. used to map between the Earth and a unit radius sphere
uses as a model of the Earth.

```{r crs}
# longlat for a spherical version of the Earth
print(fm_crs("longlat_globe"))

# longlat for a sphere of radius 1m
print(fm_crs("longlat_norm"))

# A sphere of radius 1m
print(fm_crs("sphere"))
```
