---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  dev = "png",
  dev.args = list(type = "cairo-png"),
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# fmesher: Triangle Meshes and Other Geometry Tools

<!-- badges: start -->
[![CRAN status](http://www.r-pkg.org/badges/version-last-release/fmesher)](https://cran.r-project.org/package=fmesher)
[![inlabru-org r-universe status](https://inlabru-org.r-universe.dev/badges/fmesher)](https://inlabru-org.r-universe.dev)
[![R build status](https://github.com/inlabru-org/fmesher/workflows/R-CMD-check/badge.svg)](https://github.com/inlabru-org/fmesher/actions)
[![test-coverage](https://github.com/inlabru-org/fmesher/workflows/test-coverage/badge.svg)](https://github.com/inlabru-org/fmesher/actions)
[![Codecov test coverage](https://codecov.io/gh/inlabru-org/fmesher/branch/devel/graph/badge.svg)](https://app.codecov.io/gh/inlabru-org/fmesher?branch=devel)
<!-- badges: end -->

Generate planar and spherical triangle meshes,
compute finite element calculations for 1- and 2-dimensional flat and curved
manifolds with associated basis function spaces, methods for lines and
polygons, and transparent handling of coordinate reference systems and
coordinate transformation, including `sf` and `sp` geometries.
The core `fmesher` library code was originally part of the
[`INLA`](https://www.r-inla.org) package, and also distributed in the
[EUSTACE Horizon 2020 project](https://github.com/eustace-data/eustace-ambitious),
and implements parts of "Triangulations and Applications" by
[Hjelle and Dæhlen (2006)](https://doi.org/10.1007/3-540-33261-8).
The expanded `crs`/`CRS` support started as an add-on feature of
[`inlabru`](https://inlabru-org.github.io/inlabru/).
    
## Installation

You can install the current [CRAN version](https://cran.r-project.org/package=fmesher) version of `fmesher`:
```{r cran-installation, eval = FALSE}
install.packages("fmesher")
```
or
```{r cran-installation-pak, eval = FALSE}
# install.packages("pak")
pak::pak("inlabru")
```

### Development version on r-universe 

Track the development version builds via [inlabru-org.r-universe.dev](https://inlabru-org.r-universe.dev/builds):
```{r universe-installation-pak, eval = FALSE}
options(repos = c(
  inlabruorg = "https://inlabru-org.r-universe.dev",
  getOption("repos")
))
pak::pak("fmesher")
```
This will pick the r-universe version if it is more recent than the CRAN version.

### Development version on github

Install the development version [GitHub](https://github.com/inlabru-org/inlabru) with
```{r gh-installation-pak, eval = FALSE}
pak::pak("inlabru-org/inlabru")
```

## Online documentation

[https://inlabru-org.github.io/fmesher/](https://inlabru-org.github.io/fmesher/)

## Examples

### 2D triangular meshes

Refined constrained Delaunay triangulations can be constructed by
`fm_rcdt_2d()` and
`fm_mesh_2d()`. The `_inla()` versions of these will usually return the same
meshes as the old `INLA` methods, `INLA::inla.mesh.create()` and
`INLA::inla.mesh.2d()`.
```{r example2}
suppressPackageStartupMessages(library(fmesher))
suppressPackageStartupMessages(library(ggplot2))

bnd <- fm_extensions(cbind(0, 0), convex = c(1, 1.5))
(mesh <- fm_mesh_2d(
  boundary = bnd,
  max.edge = c(0.2, 0.5)
))
```
```{r example2-plot,fig.alt="2D triangular mesh"}
ggplot() +
  geom_fm(data = mesh) +
  theme_minimal()
```

Mostly regular triangulations can be constructed by supplying a regular set of
input points. The `fm_hexagon_lattice()` function (developed by Man Ho Suen)
generates points in a regular hexagonal lattice pattern, contained in a given
`sf` polygon.
```{r example2hex}
hex_points <- fm_hexagon_lattice(bnd = bnd[[1]], edge_len = 0.2)
(mesh_hex <- fm_mesh_2d(
  loc = hex_points,
  boundary = bnd,
  max.edge = c(0.3, 0.5)
))
```

```{r example2hex-plot,fig.alt="Quasi-regular 2D triangular mesh"}
ggplot() +
  geom_fm(data = mesh_hex) +
  theme_minimal()
```


### 1D B-spline function spaces

```{r example1}
(mesh <- fm_mesh_1d(c(1, 2, 3, 4, 6),
  boundary = c("neumann", "free"),
  degree = 2
))
```
```{r example1-plot,fig.alt="1D B-spline function space"}
ggplot() +
  geom_fm(data = mesh, xlim = c(0, 7))
```

### Extended helper methods for CRS handling

The package provides methods `fm_crs()` and `fm_CRS()` for extracting CRS
information from `sf` and `sp` objects and automatically converts to the
desired output format. The `fm_transform()` wrapper similarly handles a variety
of objects, as well as special handling for converting between spheres and globes
of different radii, e.g. used to map between the Earth and a unit radius sphere
uses as a model of the Earth.

```{r crs}
# longlat for a spherical version of the Earth
print(fm_proj4string(fm_crs("longlat_globe")))

# longlat for a sphere of radius 1m
print(fm_proj4string(fm_crs("longlat_norm")))

# A sphere of radius 1m
print(fm_proj4string(fm_crs("sphere")))
```
