---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  dev = "png",
  dev.args = list(type = "cairo-png"),
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# fmesher: Triangle Meshes and Other Geometry Tools

<!-- badges: start -->
[![inlabru status badge](https://inlabru-org.r-universe.dev/badges/fmesher)](https://inlabru-org.r-universe.dev)
[![R build status](https://github.com/inlabru-org/fmesher/workflows/R-CMD-check/badge.svg)](https://github.com/inlabru-org/fmesher/actions)
[![test-coverage](https://github.com/inlabru-org/fmesher/workflows/test-coverage/badge.svg)](https://github.com/inlabru-org/fmesher/actions)
[![Codecov test coverage](https://codecov.io/gh/inlabru-org/fmesher/branch/devel/graph/badge.svg)](https://codecov.io/gh/inlabru-org/fmesher?branch=devel)
<!-- badges: end -->

Functions that generate planar and spherical triangle meshes,
computes finite element calculations for 1- and 2-dimensional flat and curved
manifolds, functions for lines and polygons, and transparent handling of
crs information and coordinate transformation, including and interoperability
with `sf` and `sp` geometries. Experimental package that will take over the
relevant subset of features from the [INLA](https://www.r-inla.org) and
[inlabru](https://inlabru-org.github.io/inlabru/) packages, as well as extend
them.
    
## Installation

You can install the latest stable release of `fmesher` from
[GitHub](https://github.com/):
``` r
# install.packages("remotes")
remotes::install_github("inlabru-org/fmesher", ref = "stable")
```

And the development version with:
``` r
# install.packages("remotes")
remotes::install_github("inlabru-org/fmesher")
```
or track the development version builds via [inlabru-org.r-universe.dev](https://inlabru-org.r-universe.dev/builds):
``` r
# Enable universe(s) by inlabru-org
options(repos = c(
  inlabruorg = "https://inlabru-org.r-universe.dev",
  INLA = "https://inla.r-inla-download.org/R/testing",
  CRAN = "https://cloud.r-project.org"
))

install.packages("fmesher")
```
Note that you need the `feature/fmesher` branch of `inlabru` until that is merged
into the main development branch:
``` r
# install.packages("remotes")
remotes::install_github("inlabru-org/inlabru@feature/fmesher")
```

To install and run `fmesher` in full debug mode (this is quite an experience!),
use
``` r
# install.packages("pkgbuild")
source("https://raw.githubusercontent.com/inlabru-org/fmesher/devel/misc/build.R")
fmesher_install(repo = "inlabru-org/fmesher", debug = TRUE)
```



## Online documentation

[https://inlabru-org.github.io/fmesher/](https://inlabru-org.github.io/fmesher/)

## Example

```{r example}
suppressPackageStartupMessages(library(fmesher))

# Port of the old inla mesh inla.mesh.create interface:
(mesh <- fm_rcdt_2d_inla(
  cbind(0, 0),
  extend = list(offset = 1, n = 16L),
  refine = list(max.edge = 0.5)
))

plot(mesh)

# longlat for a spherical version of the Earth
print(fm_crs("longlat_globe"))

# longlat for a sphere of radius 1m
print(fm_crs("longlat_norm"))

# A sphere of radius 1m
print(fm_crs("sphere"))
```
