---
title: "Converting legacy INLA mesh code to fmesher"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Converting legacy INLA mesh code to fmesher}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = FALSE}
suppressPackageStartupMessages(library(fmesher))
suppressPackageStartupMessages(library(tibble))

convert_fun_links <- function(df) {
  webref <- list(
    fmesher = "../reference/",
    inlabru = "https://inlabru-org.github.io/inlabru/reference/"
  )
  for (k in seq_len(nrow(df))) {
    df$INLA[[k]] <- paste0("`", df$INLA[[k]], "`")
    df$fmesher[[k]] <-
      vapply(
        df$fmesher[[k]],
        function(x) {
          pkg_name <- sub(
            pattern = "^(([^:]*)::)?([^(]+)(\\(.*\\))(<-)?$",
            replacement = "\\2",
            x = x
          )
          if (identical(pkg_name, "")) {
            pkg_name <- "fmesher"
          }
          nm <- sub(
            pattern = "^(([^:]*)::)?([^(]+)(\\(.*\\))(<-)?$",
            replacement = "\\3",
            x = x
          )
          args <- sub(
            pattern = "^(([^:]*)::)?([^(]+)(\\(.*\\))(<-)?$",
            replacement = "\\4\\5",
            x = x
          )
          help_name <- basename(utils::help((nm), package = (pkg_name)))
          paste0(
            '<a href="', webref[[pkg_name]], help_name, '.html">`',
            if (!identical(pkg_name, "fmesher")) {
              paste0(pkg_name, "::")
            } else {
              ""
            },
            nm,
            args,
            "`</a>"
          )
        },
        ""
      )
  }
  df
}
```

## Deprecation warnings

Three `INLA::inla.getOption()`/`INLA::inla.setOption()` options control the
transition behaviour of the `INLA` package use of `fmesher`.

* `fmesher.evolution`, integer:
  - `1L` uses the intermediate `fm_*` methods in `fmesher` that were already available via inlabru from 2.8.0,
    but calls the `INLA` built-in `fmesher` standalone programme for mesh construction
    and related operations. (From `INLA` version 23.06.29)
  - `2L` uses the full range of `fmesher` package methods, and does not use the standalone `fmesher` programme.
    (From `INLA` around version 23.08.20)

* `fmesher.evolution.warn`, logical:
  When `TRUE`, `INLA` will show deprecation methods for all the methods in `INLA` that
  are being replaced by `fmesher` package methods.  When `FALSE`, no warnings will be shown.
  Set this option to `TRUE` if you want to update your own code, but keep it at
  `FALSE` when you need to run existing code without changing it.
  
* `fmesher.evolution.verbosity`, character: Either "soft", "warn", or "stop",
  indicating the minimum warning level when `fmesher.evolution.warn` is `TRUE`.
  Set this to "warn" or "stop" to get more immediate feedback when testing
  conversion of old code, e.g. in package testing.
  
### Compatibility

Great effort has been taken to preserve backwards compatibility as far as practical,
with in particular the old `inla.mesh`, `inla.mesh.1d`, and `inla.mesh.segment`
object classes given fallback methods that carry out methods for the new `fm_mesh_2d`,
`fm_mesh_1d`, and `fm_segm` classes.  An important change is that the handling
of mesh `crs` information is now more flexible, but stricter in the sense that
user code should make no assumption about how the information is stored in the
mesh, and should therefore avoid the direct `mesh$crs` access, and instead use
the `fm_crs()` and `fm_CRS()` access methods,
depending on what type of CRS object is needed.  The ideal way to
specify crs information is in the initial mesh creation call. If the crs needs
to be explicitly assigned a new value, use the `fm_crs(mesh) <- crs` assignment
method.


## Mesh construction

```{r echo = FALSE}
df <- tribble(
  ~INLA, ~fmesher,
  "inla.mesh.create()", c("fm_rcdt_2d()", "fm_rcdt_2d_inla()"),
  "inla.mesh.2d()", c("fm_mesh_2d()", "fm_mesh_2d_inla()"),
  "inla.delaunay()", "fm_delaunay_2d()",
  "inla.mesh.1d()", "fm_mesh_1d()",
  "inla.mesh.lattice()", "fm_lattice_2d()",
  "inla.mesh.segment()", "fm_segm()",
  "inla.nonconvex.hull()", c("fm_nonconvex_hull()", "fm_extensions()", "fm_simplify()"),
  c(
    "inla.nonconvex.hull()",
    "inla.contour.segment()",
    "inla.simplify.curve()"
  ),
  c(
    "fm_nonconvex_hull_inla()",
    "fm_simplify_helper()",
    "fm_segm_contour_helper()"
  )
)
df <- convert_fun_links(df)
```

```{r, echo = FALSE}
knitr::kable(df)
```

## Basis and function evaluation

```{r echo = FALSE}
df <- tribble(
  ~INLA, ~fmesher,
  "inla.mesh.projector()", "fm_evaluator()",
  "inla.mesh.project()", "fm_evaluate()",
  "inla.spde.make.A()", c(
    "fm_basis()",
    "fm_row_kron()",
    "inlabru::bru_mapper_multi()",
    "inlabru::ibm_jacobian()",
    "fm_block()",
    "fm_block_eval()",
    "inlabru::bru_mapper_aggregate()"
  ),
  "inla.mesh.deriv()", "fm_basis()"
)
df <- convert_fun_links(df)
```

```{r, echo = FALSE}
knitr::kable(df)
```

## Finite element methods

```{r echo = FALSE}
df <- tribble(
  ~INLA, ~fmesher, ~Comment,
  c("inla.mesh.fem()", "inla.mesh.1d.fem()"), "fm_fem()", " ",
  " ", "fm_matern_precision()", " ",
  " ", "fm_matern_sample()", "Convenience function that combines `fm_matern_precision()` and `fm_sample()`.",
  " ", "fm_covariance()", "Basic helper function for compution covariances between different locations. Can produce sparse inverses like `inla.qinv()`, but currently (version 0.1.1) only by a 'brute force' method.",
  " ", "fm_sample()", "Basic sampling method, like `inla.qsample()`"
)
df <- convert_fun_links(df)
```

```{r, echo = FALSE}
knitr::kable(df)
```

## Printing

```{r echo = FALSE}
df <- tribble(
  ~INLA,
  ~fmesher,
  ~Comments,
  "summary.inla.mesh()",
  c(
    "print.fm_mesh_2d()",
    "print.fm_segm()",
    "print.fm_mesh_1d()"
  ),
  "Use `print(mesh)` etc."
)
df <- convert_fun_links(df)
```

```{r, echo = FALSE}
knitr::kable(df)
```

## CRS information and coordinate transformations

```{r, echo = FALSE}
df <- convert_fun_links(
  tibble::tribble(
    ~INLA, ~fmesher, ~Comments,
    "inla.spTransform()", "fm_transform()", "",
    "mesh$crs", c("fm_crs(mesh)", "fm_CRS(mesh)"), "The crs may now be stored in different formats; use `fm_crs()` for `sf` format, and `fm_CRS()` for `sp` format. `fmesher` will attempt to convert when needed.",
    "mesh$crs<-", "fm_crs(mesh)<-", "Direct assignment of crs information should be avoided, but is allowed as long as its compatible with the actual mesh coordinates."
  )
)
```

```{r, echo = FALSE}
knitr::kable(df)
```

## Plotting

```{r echo = FALSE}
df <- tribble(
  ~INLA, ~fmesher, ~Comment,
  "No ggplot support", c("geom_fm(data = mesh)", "geom_fm(data = segm)"), "Use `ggplot() + geom_fm(data = mesh)` and `inlabru::gg()` methods",
  "plot.inla.mesh(rgl = FALSE)", c("plot.fm_mesh_2d()", "lines.fm_mesh_2d()"), "Use `plot()` or `lines()`",
  "lines.inla.mesh.segment(rgl = FALSE)", c(
    "plot.fm_segm()",
    "lines.fm_segm()"
  ), "Use `plot()` or `lines()`",
  "plot.inla.mesh(rgl = TRUE)", c("plot_rgl()", "lines_rgl()"), "",
  "lines.inla.mesh.segment(rgl = TRUE)", c(
    "plot_rgl()",
    "lines_rgl()"
  ),
  ""
)
df <- convert_fun_links(df)
```

```{r, echo = FALSE}
knitr::kable(df)
```

